Verification of a Deterministic Random Bits Generator

### 2015-12-23 paper/paper.tex
added the entropy section

\section{Introduction}
A DRBG is a program that takes a source of randomness, or entropy, and stretches it out into a string of pseudorandom bits. \stdtitle{NIST SP 800-90A} DRBG provides backtracking resistance and, optionally, prediction resistance. Backtracking resistance means that even if the whole (secret) state of the DRBG is known at time $t$, the state at time $t - 1$ cannot be computed. Note that this implies the state at time $t - i$ cannot be computed for all $i \geq 1$. Prediction resistance means that even if the whole (secret) state of the DRBG is known at time $t$, the state at time $t + 1$ cannot be computed. Again, note that this implies the state at time $t + i$ cannot be computed for all $i \geq 1$. Here, “time” is counted in terms of calls to the \lstinline{generate} function.

@@

>>>>

\section{Introduction}
A DRBG is a program that takes a source of randomness, or entropy, and stretches it out into a string of pseudorandom bits. \stdtitle{NIST SP 800-90A} \cite{nist} DRBG provides backtracking resistance and, optionally, prediction resistance. Backtracking resistance means that even if the whole (secret) state of the DRBG is known at time $t$, the state at time $t - 1$ cannot be computed. Note that this implies the state at time $t - i$ cannot be computed for all $i \geq 1$. Prediction resistance means that even if the whole (secret) state of the DRBG is known at time $t$, the state at time $t + 1$ cannot be computed. Again, note that this implies the state at time $t + i$ cannot be computed for all $i \geq 1$. Here, “time” is counted in terms of calls to the \lstinline{generate} function.

@@

\section{A Functional Representation of Entropy} \label{entropy}
Before even understanding what a DRBG does, I need to explain how entropy is represented.

A DRBG needs some entropy input to act as a source of randomness. According to \stdtitle{NIST SP 800-90A}, the source of entropy input can be either an approved entropy source, a nondeterministic random bits generator (NRBG), or another DRBG. The entropy API used in NIST is the following:

\begin{lstlisting}
    (status, entropy_input) = Get_entropy_input(min_entropy, min_length, max_length, prediction_resistance_request)
\end{lstlisting}

\lstinline{min_entropy} refers to how many bits of entropy to use. \lstinline{min_length} and \lstinline{max_length} bounds the length of the output string of bits. If \lstinline{prediction_resistance_request} is set, then fresh bits of entropy are used.

For this project, I wanted a simple entropy representation that still satisfies \stdtitle{NIST SP 800-90A}. A flawed idea is to represent entropy as an infinite stream of \lstinline{bool}, each \lstinline{bool} representing a bit. This does not model real-life entropy. Entropy is expensive; if it was not, there would be no need for a DRBG to begin with! Requesting too many bits of entropy when there is insufficient supply can cause a failure. To capture this behavior, I chose to model entropy as an infinite stream of \lstinline{option bool}.

Note that in Coq, as well as many typed programming languages, an \lstinline{option} type is a type that can simply not exist (\lstinline{null} in the language of C and Java). It is effectively equivalent to this:

\begin{lstlisting}
    Inductive option (T: Type) :=
    | Some: T -> option T
    | None: option T.
\end{lstlisting}

A \lstinline{None} value means that the entropy bit is not yet available, and so the call to get entropy should fail. An infinite stream of entropy can be defined as follows:

\begin{lstlisting}
    Definition stream: Type := nat -> option bool.
\end{lstlisting}

The \lstinline{nat} (natural number) argument is used to index into the stream. Note that a simple \lstinline{list} would not work here, because we need the stream to have infinite size.

Any functions that consume entropy will need to return a new entropy stream. Furthermore, because calls to get entropy can fail, such functions have to be able to handle failures. For this reason, I defined a general \lstinline{result} datatype:

\begin{lstlisting}
    Inductive error_code: Type :=
    | catastrophic_error
    | generic_error.

    Inductive result (X: Type): Type :=
    | success: X -> stream -> result X
    | error : error_code -> stream -> result X.
\end{lstlisting}

I also defined a \lstinline{get_bits} function that, if possible, get the consecutive requested number of bits from the stream and returns a new stream that advances the index by the requested number of bits. If \lstinline{get_bits} encounters an error (a \lstinline{None} value), it returns a new stream that skips this error the next time it is called, symbolizing a movement forward in time resulting in more entropy being available.

This stream representation of entropy is a specialized case of the \lstinline{Get_entropy_input} function from \stdtitle{NIST SP 800-90A}, with each bit of the stream representing a bit of entropy, therefore setting \lstinline{min_entropy = min_length = max_length}, with prediction resistance always on. In the language of \stdtitle{NIST SP 800-90A}, this entropy representation is an “ideal random bitstring”.
